\mytitle

### 2020-05-18 paper.tex
speedup results and renaming.

%\newcommand{\mytitle}{Exploiting Software and Hardware Advances  in Optimization Methods for Improved Wind Farm Design}
\newcommand{\mytitle}{Exploiting Hardware and Software Advances 
 for Quadratic Models of Wind Farm Layout Optimization}
%\newcommand{\mytitlerunning}{\mytitle}
%\newcommand{\myauthor}{Arik Senderovich, Eldan Cohen, and J. Christopher Beck}
%\newcommand{\myauthorhyperref}{Arik Senderovich, Eldan Cohen, and J. Christopher Beck}

@@

To 
maximize $f(x)$ and satisfy the total number of turbines and proximity 
constraints we write
the following quadratic optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars}
\end{eqnarray} Constraints~\ref{Cardinality}~and~\ref{Grid} ensure that exactly $m$ turbines are placed on the grid
and enforce proximity constraints between the turbines, respectively. We shall refer to the QCOP
formulation of WFLO as WFL-QCOP.

WFL-QCOP can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
%\todo{JCB: We need to formally present the ILP model in the experiment section. We can then discuss the fact that the model is identical - because the linearization gives and exact representation. Should put a forward reference to that model here.}
%\todo{Are these linear formulations ``equivalent'' as you claimed or approximations? If they are mathematically equivalent, why do we think a quadratic model will be any better computationally? Arik: The two formulations are equivalent to WFL-QCOP. However, we compare against the ILP model experimentally and it does worse than WFL-QCOP. Hopefully, Jason's part in Section 3.5.1. based on Wen-Yang's thesis will explain why it does better. Originally, I used the quadratic formulation because it seemed more natural. Semantically, the linear model and the quadratic model are the equivalent imho.}

@@

Let $\lambda = (\lambda_1,\lambda_2)$ 
be a vector of constraint parameters with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq. (\ref{QCOP}) is given by,
\begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
penalizing solutions that violate the exactly $m$ turbines constraint and the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizing
solutions that place turbines too close together. The two penalty terms $\lambda_1$ and $\lambda_2$ 
must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints. We shall refer to the QUBO representation of WFLO as WFL-QUBO.
%These types of constraints
%are referred to as \emph{soft constraints} as they do not guarantee a feasible solution 
%for an arbitrary selection of $\lambda$.

\subsection{The Computational Complexity of WFL-QCOP}
\label{sec:computational}
%(linear) problems were formulated and solved using  exact solvers.
Despite the existence of mathematical models for WFLO in the literature~\cite{Zhang2014,donovan2005wind} that are expressive enough to represent $\mathcal{NP}$-hard problems, we are unaware of an explicit treatment of the computational complexity of WFLO.
%However, these solvers should only be used if the problem is computationally hard.
%We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP).
We therefore provide a straightforward proof of the computational complexity of WFL-QCOP.

\begin{mythm}
	The computational complexity of WFL-QCOP defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
\end{mythm}
\begin{myproof}
We prove by showing that a special case of the WFL-QCOP \eqref{QCOP}-\eqref{Vars}  is
$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
Since a special case of WFL-QCOP is  $\mathcal{NP}$-hard it implies that WFL-QCOP is at least $\mathcal{NP}$-hard. On the other hand,
WFL-QCOP is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that WFL-QCOP is $\mathcal{NP}$-hard.
\end{myproof}

@@

\subsubsection{QCOP solving methods}
\label{quadratic}
WFL-QCOP is a strictly convex integer 
quadratically constrained optimization problem (IQCOP), which is notorious for its computational complexity~\cite{van1981another}.
%In fact, we proved in Section~\ref{sec:computational} that WFL-QCOP is $\mathcal{NP}$-hard.
The common exact methods for solving
IQCOPs are tree search \cite{land2010automatic} algorithms with a variety of sophisticated techniques to limit the search necessary to find and prove optimality.  In operations research, the generic exact approach for solving IQCOPs is the branch-and-cut based mixed integer programming (MIP) \cite{bonami2008algorithmic}. A MIP solver, such as the one we use (Gurobi), uses a linear or quadratic continuous relaxation to efficiently compute a bound on the objective function. This relaxation is used to generate cuts, implied constraints that tighten the relaxation, to identify sub-trees that can be discarded without search because they are guaranteed not to contain a solution better than the current incumbent, and to drive the heuristic branching that systematically searches for a solution. Typically, this process of cut generation, reasoning about sub-trees, and branching is performed at each node in the search tree.

Commercial MIP solvers have seen extensive development over the past 20 years and hardware-independent algorithmic improvements delivering orders of magnitude improved performance \cite{Bixby07a}. More recently the development of such solvers has focused on improved performance for quadratic problems such as WFLO-QCOP \cite{Furini19a}.
%By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.

@@

In this section, we 
present an empirical evaluation of the two quadratic models based on 
12 standard WFLO instances.
Our main results are: \begin{itemize}
	\item WFL-QUBO when solved using the digital annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
		\item To achieve such a solution quality, the runtime for solving WFL-QUBO on the DA is at least two orders of magnitude lower than WFL-QCOP for hard WFLO instances.  
	\item Given sufficient runtime, solving the WFL-QCOP model yields strong results on most of the instances. Compared to previous exact models (QP-SS and ILP-LS), WFL-QCOP is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard scenarios. %  \todo{Arik: Is this correct? Table 2 shows DA@10seconds with the best solution found for 7/12 instances and QP-LS@3600 for 6/12. And 4/6 vs. 3/6 for the WR36). Are you comparing QP-LS with QP-SS and ILP-LS? We seem to be underplaying DA performance in this comment. Chris: I changed to competitive results. You're right, I'd say that the DA is not `losing' to QP-LS.}  
\end{itemize}

@@

We compared our two models, WFL-QCOP and WFL-QUBO, against two state-of-the-art exact optimization methods:
the integer linear programming (ILP-LS) approach~\cite{donovan2005wind,Zhang2014} and the 
model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. 
The ILP model is a standard, exact linearization of the WFL-QCOP model.
%% given in the following set of equations, which we denote WFL-ILP, \begin{eqnarray} \label{ILP}
%% &\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)y_{i,j})   \\
%% &s.t.:& \nonumber\\
%% &\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality2}\\
%% &\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid2}\\
%% &\mbox{       }& x_i + x_j -1 \leq y_{i,j},  \forall i,j \in \mathcal{N}, \label{Linearization}\\
%% &\mbox{       }& y_{i,j} \geq 0,     \forall i,j \in \mathcal{N}, \label{Yij}\\ 
%% &\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars2} 
%% \end{eqnarray} The variables $y_{i,j}$ are continuous and can take only non negative values (Eq.(\ref{Yij})). However, due to Eq.~(\ref{Yij}) their values must
%% be either zero or one. If a turbine is placed in both $i$ and $j$ ($x_i = x_j = 1$), then $y_{i,j}$ will be $1$ and a wake reduction 
%% will take place in the objective function. If a turbine is placed in either $x_i$ or $x_j$ (but not in both), $y_{i,j}$ will
%% take the value of $0$ and no wake reduction will occur. The WFL-ILP is equivalent to WFL-QCOP: the interaction
%% variables $y_{i,j}$ of the ILP model are replaced by multiplicative expressions $x_{i}x_{j}$, which makes Constraint (\ref{Yij}) redundant. 
However, the ILP is solved using linear optimization approaches, while WFL-QCOP is solved using quadratic programming methods as mentioned in Section~\ref{quadratic}. Hence
we can expect a difference in the performance of optimization software when solving the two problems.

The existing constraint programming approach was
reported to perform worse than the ILP model~\cite{Zhang2014} and so was omitted from our the experiments.
We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-LS and QP-SS. 
For consistency, we shall refer to WFL-QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
In addition, we 
used both Fujitsu's Digital Annealer and Gurobi to solve WFL-QUBO: we shall refer to the former as DA-LS, and to the latter
as QUBO-LS. \todo{Arik: I think it is important to use consistent names through-out. The problem is that we do not have a unified naming format. The problems can be QC-LS, QU-LS (for Quadratic Constraint/Unconstrained Linear Superposition) instead of WFL-QCOP QFL-QUBO. Then the results need to include the solver and so we have QU-LS(DA), QU-LS(GRB), QC-LS(GRB), QC-SS(GRB), and LC-LS(GRB). (LC = Linear Constrained). I suggest you use Latex macros so that we can easily change then all again if necessary.}

@@

For exact methods (QP-SS, ILP-LS, QP-LS and QUBO-LS),
we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
found until that point in time. The exact methods were all solved using Gurobi's quadratic programming solver on a 
Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.

@@

%The best solution in terms of energy is the solution that attained the highest objective
%values (SS in QP-SS and LS in the rest of the methods).

@@

We measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss})).
%\footnote{The total mean energy is presented in kW.}.
The only method that directly optimizes SS is QP-SS. Hence, for the other methods, we converted solutions
LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}. Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
Then, the SS energy expression can be written as, 
$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$

@@

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different methods.
%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \mbox{
	\{\emph{QUBO-LS}}, \mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, \mbox{\emph{ILP-LS}}, \mbox{\emph{QP-SS}}\}$. Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the relative error at time $t$ for approach $a$ on instance $w$ is given by
\begin{equation}
  RE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.
\end{equation} Note that the expression is always positive, since we are solving a maximization problem. We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} RE(w,t,a).\end{equation}
We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.

The last measure that we introduce is the speed-up factor $S(a_1, a_2, t, w)$. The measure compares the time that it takes
approach $a_1 \in \mathcal{A}$ to
achieve the same energy level as the energy achieved by approach $a_2 \in \mathcal{A}$
within $t$ seconds when solving instance $w$ (energy levels are measured in terms of the sum-of-squares energy expression). 
Formally, $$S(a_1, a_2, t, w) = \frac{\min \big[\{ t' : E_{SS}(a_1, t',w) = E_{SS}(a_2, t, w) \} \cup \{3600\} \big] }{t}, $$
with $E_{SS}(a, t, w)$ being the SS energy value under approach $a$ for instance $w$ at time $t$. The speed-up factor 
is useful to measure how many times faster (or slower) approach $a_1$ solves problem $w$ compared to approach $a_2$.   
For example, $$S(\mbox{\emph{QP-LS}}, \mbox{\emph{DA-LS}}, 10, w_1)$$ is the ratio between the time that it takes QP-LS to 
reach the same quality of solution as DA-LS achieves after $10$ seconds on problem $w_1$. Since
we limit the runtime of the exact methods by $3600$ seconds, $S$ is a lower bound on the true speed-up factor. 
%Therefore, if for the example above the result is $3600/10$, we shall interpret it as a speed-up factor of \emph{at least} $360$.

@@

%% In this part, we present the main findings of our empirical evaluation. 
%% We shall start by showing the best attained solutions by all methods.
Table~\ref{tab:results1} contains the best energy (in kW) attained by the methods
for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. 
The runtime for DA-LS was always less than $10$ seconds.

>>>>

%\newcommand{\mytitle}{Exploiting Software and Hardware Advances  in Optimization Methods for Improved Wind Farm Design}
\newcommand{\mytitle}{Exploiting Hardware and Software Advances 
 for Quadratic Models of Wind Farm Layout Optimization}
%\newcommand{\mytitlerunning}{\mytitle}
%\newcommand{\myauthor}{Arik Senderovich, Eldan Cohen, and J. Christopher Beck}
%\newcommand{\myauthorhyperref}{Arik Senderovich, Eldan Cohen, and J. Christopher Beck}
\newcommand{\qcls}{{\sf {\small QC-LS\xspace}}}
\newcommand{\ilpls}{{\sf {\small ILP-LS\xspace}}}
\newcommand{\quls}{{\sf {\small QU-LS\xspace}}}
\newcommand{\qcss}{{\sf {\small QC-SS\xspace}}}

\newcommand{\qclsgrb}{{\sf {\small QC-LS(GRB)\xspace}}}
\newcommand{\ilplsgrb}{{\sf {\small ILP-LS(GRB)\xspace}}}
\newcommand{\qulsgrb}{{\sf {\small QU-LS(GRB)\xspace}}}
\newcommand{\qcssgrb}{{\sf {\small QC-SS(GRB)\xspace}}}
\newcommand{\qulsda}{{\sf {\small QU-LS(DA)\xspace}}}

@@

To 
maximize $f(x)$ and satisfy the total number of turbines and proximity 
constraints we write
the following quadratic optimization problem (QCOP):
\begin{eqnarray} \label{QCOP}
&\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)x_i x_j)   \\
&s.t.:& \nonumber\\
&\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality}\\
&\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid}\\
&\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars}
\end{eqnarray} Constraints~\ref{Cardinality}~and~\ref{Grid} ensure that exactly $m$ turbines are placed on the grid
and enforce proximity constraints between the turbines, respectively. We shall refer to the QCOP
formulation of WFLO as \qcls{}.

\qcls{} can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
%\todo{JCB: We need to formally present the ILP model in the experiment section. We can then discuss the fact that the model is identical - because the linearization gives and exact representation. Should put a forward reference to that model here.}
%\todo{Are these linear formulations ``equivalent'' as you claimed or approximations? If they are mathematically equivalent, why do we think a quadratic model will be any better computationally? Arik: The two formulations are equivalent to WFL-QCOP. However, we compare against the ILP model experimentally and it does worse than WFL-QCOP. Hopefully, Jason's part in Section 3.5.1. based on Wen-Yang's thesis will explain why it does better. Originally, I used the quadratic formulation because it seemed more natural. Semantically, the linear model and the quadratic model are the equivalent imho.}

@@

Let $\lambda = (\lambda_1,\lambda_2)$ 
be a vector of constraint parameters with $\lambda_i >0$. The equivalent QUBO formulation of the QCOP in Eq. (\ref{QCOP}) is given by,
\begin{equation}\max_{x}^{} f(x) - \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2 - \lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j , \label{QUBO}\end{equation}
with the term $ \lambda_1 (\sum_{i \in \mathcal{N}}^{} x_i -m) ^2$ 
penalizing solutions that violate the exactly $m$ turbines constraint and the term $\lambda_2 \sum_{(i,j) \in \mathcal{E}}^{} x_i x_j$ penalizing
solutions that place turbines too close together. The two penalty terms $\lambda_1$ and $\lambda_2$ 
must be large enough to ensure that the solutions are indeed feasible wrt the problem constraints. We shall refer to the QUBO representation of WFLO as \quls.
%These types of constraints
%are referred to as \emph{soft constraints} as they do not guarantee a feasible solution 
%for an arbitrary selection of $\lambda$.

\subsection{The Computational Complexity of \qcls{}}
\label{sec:computational}
%(linear) problems were formulated and solved using  exact solvers.
Despite the existence of mathematical models for WFLO in the literature~\cite{Zhang2014,donovan2005wind} that are expressive enough to represent $\mathcal{NP}$-hard problems, we are unaware of an explicit treatment of the computational complexity of WFLO.
%However, these solvers should only be used if the problem is computationally hard.
%We are unaware of an existing complexity result for the LS variation of WFLO (our QCOP).
We therefore provide a straightforward proof of the computational complexity of \qcls{}.

\begin{mythm}
	The computational complexity of \qcls{} defined in \eqref{QCOP}-\eqref{Vars} is $\mathcal{NP}$-hard. 
\end{mythm}
\begin{myproof}
We prove by showing that a special case of the \qcls{} \eqref{QCOP}-\eqref{Vars}  is
$\mathcal{NP}$-hard. Suppose that $\mathcal{E} = \emptyset$, i.e., the grid is coarse grained enough to ignore proximity constraints. Then, the constraints in Eq.~(\ref{Grid})
are dropped and we get the heaviest k-subgraph problem, which was proven to be $\mathcal{NP}$-hard~\cite{billionnet2005different}.
Since a special case of \qcls{} is  $\mathcal{NP}$-hard it implies that \qcls{} is at least $\mathcal{NP}$-hard. On the other hand,
\qcls{} is formulated using integer programming (Problem \ref{QCOP}), which means that it is at most $\mathcal{NP}$-hard. Therefore, we get that \qcls{} is $\mathcal{NP}$-hard.
\end{myproof}

@@

\subsubsection{QCOP solving methods}
\label{quadratic}
\qcls{} is a strictly convex integer 
quadratically constrained optimization problem (IQCOP), which is notorious for its computational complexity~\cite{van1981another}.
%In fact, we proved in Section~\ref{sec:computational} that WFL-QCOP is $\mathcal{NP}$-hard.
The common exact methods for solving
IQCOPs are tree search \cite{land2010automatic} algorithms with a variety of sophisticated techniques to limit the search necessary to find and prove optimality.  In operations research, the generic exact approach for solving IQCOPs is the branch-and-cut based mixed integer programming (MIP) \cite{bonami2008algorithmic}. A MIP solver, such as the one we use (Gurobi), uses a linear or quadratic continuous relaxation to efficiently compute a bound on the objective function. This relaxation is used to generate cuts, implied constraints that tighten the relaxation, to identify sub-trees that can be discarded without search because they are guaranteed not to contain a solution better than the current incumbent, and to drive the heuristic branching that systematically searches for a solution. Typically, this process of cut generation, reasoning about sub-trees, and branching is performed at each node in the search tree.

Commercial MIP solvers have seen extensive development over the past 20 years and hardware-independent algorithmic improvements delivering orders of magnitude improved performance \cite{Bixby07a}. More recently the development of such solvers has focused on improved performance for quadratic problems such as \qcls{} \cite{Furini19a}.
%By solving a relaxed problem to bound the objective value at each node and strengthening the relaxed problem with cutting planes to restrict the feasible region further, MIP solvers has been proved effective for attacking IQCOPs \cite{junger200950}.

@@

In this section, we 
present an empirical evaluation of the two quadratic models based on 
12 standard WFLO instances.
Our main results are: \begin{itemize}
	\item \quls{} when solved using the digital annealer (DA) achieves a solution quality as good or better than the state-of-the-art exact methods on 10 of 12 instances.
		\item To achieve such a solution quality, the runtime for solving \quls{} on the DA is at least two orders of magnitude lower than \qcls{} for hard WFLO instances.  
	\item Given sufficient runtime, solving the \qcls{} model yields strong results on most of the instances. Compared to previous exact models (such as \qcss), \qcls{} is able to find equal or better solutions on 8 of 12 benchmark problems and for 5 of the 6 hard scenarios. %  \todo{Arik: Is this correct? Table 2 shows DA@10seconds with the best solution found for 7/12 instances and \qcls{}grb@3600 for 6/12. And 4/6 vs. 3/6 for the WR36). Are you comparing \qcls{}grb with \qcssgrb and ILP-LS? We seem to be underplaying DA performance in this comment. Chris: I changed to competitive results. You're right, I'd say that the DA is not `losing' to \qcls{}grb.}  
\end{itemize}

@@

We compared our two models, \qcls{} and \quls{}, against two state-of-the-art exact optimization methods:
the integer linear programming (denoted \ilpls) approach~\cite{donovan2005wind,Zhang2014} and the 
model that uses a quadratic approximation (\qcss)~\cite{turner2014new}. 
The ILP model is a standard, exact linearization of the \qcls{} model.
%% given in the following set of equations, which we denote WFL-ILP, \begin{eqnarray} \label{ILP}
%% &\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)y_{i,j})   \\
%% &s.t.:& \nonumber\\
%% &\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality2}\\
%% &\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid2}\\
%% &\mbox{       }& x_i + x_j -1 \leq y_{i,j},  \forall i,j \in \mathcal{N}, \label{Linearization}\\
%% &\mbox{       }& y_{i,j} \geq 0,     \forall i,j \in \mathcal{N}, \label{Yij}\\ 
%% &\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars2} 
%% \end{eqnarray} The variables $y_{i,j}$ are continuous and can take only non negative values (Eq.(\ref{Yij})). However, due to Eq.~(\ref{Yij}) their values must
%% be either zero or one. If a turbine is placed in both $i$ and $j$ ($x_i = x_j = 1$), then $y_{i,j}$ will be $1$ and a wake reduction 
%% will take place in the objective function. If a turbine is placed in either $x_i$ or $x_j$ (but not in both), $y_{i,j}$ will
%% take the value of $0$ and no wake reduction will occur. The WFL-ILP is equivalent to WFL-QCOP: the interaction
%% variables $y_{i,j}$ of the ILP model are replaced by multiplicative expressions $x_{i}x_{j}$, which makes Constraint (\ref{Yij}) redundant. 
However, the ILP is solved using linear optimization approaches, while \qcls{} is solved using quadratic programming methods as mentioned in Section~\ref{quadratic}. Hence
we can expect a difference in the performance of optimization software when solving the two problems.

The existing constraint programming approach was
reported to perform worse than the ILP model~\cite{Zhang2014} and so was omitted from our the experiments.
We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve \qcls{}, \ilpls{} and \qcss. 
We shall denote the problems using their names and the solvers used as follows: \qclsgrb{} (GRB stands for Gurobi), \ilplsgrb{}, and \qcssgrb{}.
In addition, we 
used both Fujitsu's Digital Annealer and Gurobi to solve \quls: we shall refer to the former as \qulsda{}, and to the latter
as \qulsgrb.

@@

For exact methods (\qcss{}, \ilpls{}, \qcls{} and \quls{}),
we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
found until that point in time. The exact methods were all solved using Gurobi's quadratic programming solver on a 
Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM.

@@

%The best solution in terms of energy is the solution that attained the highest objective
%values (SS in \qcssgrb and LS in the rest of the methods).

@@

We measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss})).
%\footnote{The total mean energy is presented in kW.}.
The only method that directly optimizes SS is \qcssgrb{}. Hence, for the other methods, we converted solutions
LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}. Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
Then, the SS energy expression can be written as, 
$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$

@@

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different methods.
%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by \qulsgrb and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $$\mathcal{A}= \mbox{
	\{\emph{\qulsgrb}}, \mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, \mbox{\emph{\ilplsgrb}}, \mbox{\emph{\qcssgrb}}\}.$$ Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the relative error at time $t$ for approach $a$ on instance $w$ is given by
\begin{equation}
  RE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.
\end{equation} Note that the expression is always positive, since we are solving a maximization problem. We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} RE(w,t,a).\end{equation}
We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.

The last measure that we introduce is the speed-up factor $S(a_1, a_2, t, w)$. The measure compares the time that it takes
approach $a_1 \in \mathcal{A}$ to
achieve the same energy level as the energy achieved by approach $a_2 \in \mathcal{A}$
within $t$ seconds when solving instance $w$ (energy levels are measured in terms of the sum-of-squares energy expression). 
Formally, \begin{equation} S(a_1, a_2, t, w) = \frac{\min \big[\{ t' : E_{SS}(a_1, t',w) = E_{SS}(a_2, t, w) \} \cup \{3600\} \big] }{t}, \label{eq:speedup} \end{equation}
with $E_{SS}(a, t, w)$ being the SS energy value under approach $a$ for instance $w$ at time $t$. The speed-up factor 
is useful to measure how many times faster (or slower) approach $a_1$ solves problem $w$ compared to approach $a_2$.   
For example, $$S(\mbox{\emph{\qclsgrb}}, \mbox{\emph{\qulsda}}, 10, w_1)$$ is the ratio between the time that it takes \qclsgrb{} to 
reach the same quality of solution as \qulsda{} achieves after $10$ seconds on problem $w_1$. Since
we limit the runtime of the exact methods by $3600$ seconds, $S$ is a lower bound on the true speed-up factor. 
%Therefore, if for the example above the result is $3600/10$, we shall interpret it as a speed-up factor of \emph{at least} $360$.

@@

%% In this part, we present the main findings of our empirical evaluation. 
%% We shall start by showing the best attained solutions by all methods.
Table~\ref{tab:results1} contains the best energy (in kW) attained by the methods
for the 12 WFLO instances after $100$ seconds of runtime for the exact methods. 
The runtime for the DA was always less than $10$ seconds.

\begin{table}[t!]
	\small
	\begin{tabular}{| c | c | c | c | c | c | c | c |}
		\toprule
		Wind Directions  & n  & m  & \qulsda{}  & \qulsgrb  & \qcls & \qcss  & \ilpls  \\
		\toprule
		\multirow{6}{*}{WR1}  & \multirow{3}{*}{10}       & 20       & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} & \textbf{11185.41} \\
		& & 30   & 15740.69 & \textbf{15742.93}  & 15738.45  & 15735.10  & 15739.57     \\
		& & 40 & \textbf{19265.21} & 18952.97 & \textbf{19265.21}  & \textbf{19265.21} & \textbf{19265.21}                \\
		\cline{2-8}
		&\multirow{3}{*}{20}   & 20       & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}  & \textbf{11404.80}          \\
		&&30   & 16570.96 & 16715.19  & \textbf{16758.45}  & 16755.75 & 16695.05                  \\
		&&40   & 21874.04 & 21643.56  & 21905.50 & \textbf{21977.17} & \textbf{21977.17}        \\
		\hline
		\multirow{6}{*}{WR36} &  \multirow{3}{*}{10}    & 20       & \textbf{19221.44} & \textbf{19221.44} & \textbf{19221.44} & 19193.42 & 19022.77 \\
		&& 30  & \textbf{27450.99} & 27418.74 & 27442.90 & 27387.06 & 27247.50                     \\
		&&40   & 34888.81 & \textbf{34938.65} & \textbf{34938.65} & 34817.27  & 34671.62          \\
		\cline{2-8}
		&  \multirow{3}{*}{20}   & 20   & \textbf{19441.17}  & 19172.56 & 19336.33 & 19354.23  & 18448.64            \\
		&&30   & \textbf{27975.55} & 27443.80  & 27752.00 & 27492.11  & 26367.20                      \\
		&&40   & 35604.45 & 35127.32 & \textbf{35665.60}   & 35065.85 & 33451.94 \\
		\bottomrule                   
	\end{tabular}

	\vspace{0.5em}
	\caption{Total Expected Power (kW) after 100 seconds for the exact techniques and by 10 seconds for \qulsda{}.}\label{tab:results1}
\end{table}
